Mean temperature, fluctuation magnitude, and level of biological organization affect biological responses amidst thermal variability: a meta-analysis

### 2021-06-22 thesis writing/thesis_draft/_book/thesis.tex
reorganizing repo and creating data vis file for manuscript data analysis

% Syntax highlighting #22
  \usepackage{color}
  \usepackage{fancyvrb}
  \newcommand{\VerbBar}{|}
  \newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
  \DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
  % Add ',fontsize=\small' for more characters per line
  \usepackage{framed}
  \definecolor{shadecolor}{RGB}{248,248,248}
  \newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
  \newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
  \newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
  \newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
  \newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
  \newcommand{\BuiltInTok}[1]{#1}
  \newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
  \newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
  \newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
  \newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
  \newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
  \newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
  \newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
  \newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
  \newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
  \newcommand{\ExtensionTok}[1]{#1}
  \newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
  \newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
  \newcommand{\ImportTok}[1]{#1}
  \newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
  \newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
  \newcommand{\NormalTok}[1]{#1}
  \newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
  \newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
  \newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
  \newcommand{\RegionMarkerTok}[1]{#1}
  \newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
  \newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
  \newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
  \newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
  \newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
  \newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}

@@

It is no secret that life on earth has evolved and persisted in the face of environmental change for many millennia (Bambach, 1993; Bouchard, 2014). Cyanobacteria can trace their origin back nearly 2.5 billion years to a bacteria that evolved the ability to perform oxygenic photosynthesis and changed organismal composition forever (Blankenship, 2017; Soo, Hemp, Parks, Fischer, \& Hugenholtz, 2017). Cyanobacteria have continued to persist in the world, despite experiencing widely different conditions throughout time (Knoll, 2008). Vascular plants have also evolved different photosynthetic systems (e.g.~C4, CAM) to manage a variety of environmental conditions (Guralnick, Cline, Smith, \& Sage, 2007). Though cyanobacteria and vascular plants evolved new strategies and structures to persist as the environment changed, the environment has continued to remain dynamic, such that evolutionary responses and plasticity have become even more relevant (Burggren, 2018; Reed, Waples, Schindler, Hard, \& Kinnison, 2010). Understanding how organisms respond to dynamic environmental conditions is significant for how we recapitulate ecosystems currently and in the future.

@@

Organisms and the ecosystems they inhabit have experienced variation on many scales, including diurnal changes in light availability, seasonal changes in temperature, and decadal changes in weather patterns (e.g.~El Niño Southern Oscillation). As such, the scientific literature has continued to highlight the importance of ecologically relevant studies for drawing robust conclusions about the dynamics of ecosystems (Khelifa, Blanckenhorn, Roy, Rohner, \& Mahdjoub, 2019; Stewart et al., 2013). How variability impacts populations may impact biodiversity, such that only certain species may survive, and how that biodiversity is maintained are key components to the full picture of ecosystem dynamics. As environmental patterns continue to change, from increased mean temperatures to more severe weather events (e.g.~hurricanes), understanding the limits to organisms' ability to cope with changes in environmental conditions is increasingly important for both individual level responses as well as how individual level impacts scale across population and community level responses (Bernardo, 2014; Helmuth et al., 2010; Huey et al., 2012).

@@

There is a large body of literature on environmental variability including its relevance for positive effects on populations and communities, e.g.~inflationary effects in sink-source systems (Holt, Barfield, \& Gonzalez, 2003; Vasseur \& Fox, 2009), such that variability aids population density when organisms emigrate. However, there is a lack of consensus as to which scales of biological organization, i.e.~individuals, populations, or communities of organisms, will be most severely affected by different shifts in environmental conditions. We aim to better understand how environmental variability affects the magnitude of responses across organization levels as well as important factors implicated in mediating this response. To accomplish this, we review recent progress in the field of environmental variability, specifically changes in temperature, to identify gaps between theory, experiment, and implementation.

@@

\hypertarget{environmental-variability}{%
\section{Environmental variability}\label{environmental-variability}}
\begin{figure}

{\centering \includegraphics[width=1.05\linewidth]{figures/figure1}

\caption[Metrics to describe and explain environmental variability.]{An example of environmental variability using diurnal light availability to model how range, mean, predictability, and duration can be used to characterize different patterns of variability. Shown is a comparison of how duration of variability compares within the same environmental factor. Terms highlighted in each panel are described below in the main text.}\label{fig:unnamed-chunk-3}
\end{figure}
Environmental variability describes the dynamic characteristics of abiotic and biotic environmental conditions, in addition to how abiotic and biotic factors interact to drive organisms' response to environmental conditions (Gudmundson, Eklöf, \& Wennergren, 2015; Parepa, Fischer, \& Bossdorf, 2013). Though there are several characterized patterns of environmental variability (sinusoidal, stochastic, colored noise, reviewed below), patterns of variation can be further understood using the same core set of attributes: range, mean, duration, and predictability (Figure 1.1).

@@

Aggregating values together, whether it be temperature, light availability, or precipitation, allows for a better understanding of trends and patterns as opposed to random events. The mean can be used as a zeroing point for any environmental condition that we can use to compare across patterns, studies, and time. It provides a departure point for what we expect the average value to be for all time. In certain cases, the mean is a useful metric for observing trends by minimizing the noise of environmental conditions, e.g.~averaging population growth rates to understand if a disease is affecting organisms across geographic locations or just specific locations. However, in other cases, the mean is less useful compared to the variability. For instance, if there are key factors contributing to population growth rates from specific locations, comparing the amount of variability in additional variables across locations (individual respiration rates, temperature, etc.) is more fruitful for the task at this scale. At some point however, all data of interest are averages. The important distinction is the scale at which we average the data and the level of detail provided at each scale. Means are more useful in describing variation amongst normally distributed data, however, other statistical metrics like the median or mode may be more appropriate when the data is not normally distributed.

@@

We can expand additional information adjacent to the mean (e.g.~range, duration, predictability) for a finer scale picture of environmental conditions. Vasseur et al outlined that increases in temperature variation posed a greater risk to species than just simply considering increases in global temperature (2014). We can further quantify how variable an environment is by accounting for the range, or total span of conditions an organism may experience. This range can be short or long, for example, geographic location drastically influences the amount of sunlight an ecosystem receives. The Amazon Rainforest experiences less variation in light availability during the year than Denali National Park in Alaska, which experiences high variation in light availability between the summer and winter seasons. The range of environmental conditions an organism experiences can influence their ability to withstand changes in mean temperature (Amarasekare \& Coutinho, 2013; Amarasekare \& Savage, 2012), as well as additional life history traits (size, speed, etc.). An organism's ability to withstand changes in mean temperature has been attributed to the location of their optimal thermal temperature relative to the temperature at which survival is no longer possible and may be significantly impacted by increases in temperature range (Amarasekare \& Savage, 2012). If an organism has an optimal temperature close to their upper limit of survival, a larger range of variation may push the organism to spend more time beyond the optimal temperature. This would negatively impact their ability to adapt and withstand an upward thermal shift that occurs on a faster timescale than their adaptive evolution can manage.

@@

The amount of time an organism spends at different conditions is also important for how we understand and describe environmental variability. Variability can be examined on many different scales, from daily changes in light availability to seasonal changes in light availability (Figure 1.1). Short term fluctuations, such as Fluctuating Thermal Regimes (FTR), in which an organism only experiences a change in temperature for a short period of time (less than one generation), have differing effects on performance (Colinet, Rinehart, Yocum, \& Greenlee, 2018). Performance is a set of whole organism traits (e.g.~assimilation, development, metabolic rate) with the availability to perform dynamic actions and tasks pertinent to fitness or organismal success and is often used as a way to estimate the culmination of organismal processes that hinder or help an organisms' persistence in the environment (Amarasekare \& Savage, 2012; Husak \& Lailvaux, 2017). Studies that employ FTR are doing so in an explicitly ecologically irrelevant context, as organisms reared in a cold environment do not usually experience multiple warming exposures during a single day. However, there are instances in nature where organisms do experience short periods of extreme temperature changes (e.g.~heat waves, cold fronts, etc.) in which the benefits of FTR in mitigating chill injuries may be helpful. Beyond the period of fluctuation organisms experience, it is important to also consider how the period of fluctuation aligns with individual, population, and community level processes. For instance, Temperature-dependent Sex Determination (TSD), in which ectotherms rely on the environment to incubate their larval offspring, is greatly affected by temperature changes during a specific window of development (Bowden \& Paitz, 2018; Delmas, Prevot-Julliard, Pieau, \& Girondot, 2007). Further, at the population level, if organisms are closer or farther to carrying capacity, the impact of fluctuation may help or hinder population growth depending on the magnitude of variance (Lawson, Vindenes, Bailey, \& Pol, 2015). When considering community dynamics, similar effects may appear depending on when a fluctuation occurs in time relative to predator/prey dynamics (Bastille-Rousseau et al., 2018; Dobramysl \& Tauber, 2013; Romanuk \& Kolasa, 2002).

@@

FTR can also describe the predictability of the fluctuation period. Variability may not have evenly spaced intervals (e.g.~heat waves that occur at irregular intervals within a summer season), such that organisms can experience different amounts of variability for different amounts of time. Predictability can also attempt to describe stochastic vs.~autocorrelated variation, however, that will be discussed in a different context below (Figure 1.3).

@@

While not explicitly an area of study for this work, it is important to include how temporal and spatial variability can interact. Spatial variability describes how in three dimensional space, environments, organisms, and conditions can vary based on their location in space (Farhang-Sardroodi, Darooneh, Kohandel, \& Komarova, 2019; Horne \& Schneider, 1995). Spatial variability can interact with temporal variability in cases where organisms mitigate temporally variable conditions by modifying their spatial location, e.g.~Diel vertical migration (DVM). In DVM, organisms migrate to the surface and bottom of lakes and oceans based mainly on cues of light availability (Brierley, 2014; Lampert, 1989). However, there are also instances in which the environment is not temporally variable, yet an organism utilizes spatial variability to regulate its behavior, e.g.~an organism moving from sun to shade during the day to regulate body temperature and metabolic processes (Huey, 1974; Kearney, Shine, \& Porter, 2009).

Though environmental variability can include many different forms, from fluctuations in resource availability
(Dempster \& Pollard, 1981; Sommer, 1984) to light (Morison, Franzè, Harvey, \& Menden‐Deuer, 2020; Ruel \& Ayres, 1999), one of the best studied and all-encompassing forms of fluctuations are thermal fluctuations (Colinet, Sinclair, Vernon, \& Renault, 2015; Huey \& Bennett, 1990).

@@

Changes in temperature can be grouped into a couple different definitions, the first being the difference between the terms ``fluctuating'' and ``variation'' with respect to temperature. Though many studies have implicitly assumed the audience understands the distinction between the two terms, Colinet et al (2018) explicitly defined fluctuating temperature as ``a generic term that refers to any discontinuous thermal regime that occurs on a short-term basis.'' Most generally, temperature variability describes changes in environmental temperature through time (Colinet et al., 2018; Holmes, Woollings, Hawkins, \& Vries, 2016).

@@

Most organisms experience and rely on temperature to regulate metabolic processes or signal a behavior or response to ultimately persist in the environment (Huey \& Bennett, 1990). Brown et al (2004) highlighted the importance of both temperature and body size in the metabolic theory of ecology (MTE). Though highly debated (Clarke, 2006), MTE serves as an origin point for understanding how organisms respond to environmental conditions. MTE is described by the following equation:

\[
B = b_0M^{3/4}e^{-E/kT}
\]

Where \(B\) is organismal metabolic rate, \(b_0\) is a normalization constant unrelated to temperate or body size, \(M\) is organismal mass, and \(e^-E/kT\) is referred to as the ``Boltzman factor'', where \(E\) is activation energy, \(k\) is the Boltzman constant and \(T\) is temperature. Together, all of these values aim to describe the relationship between temperature and body size to determine metabolic rate. We assume that organismal level responses should scale across higher levels of organization, as MTE predicts that after correcting for body mass and temperature, metabolic rates for larger organisms should be higher and for smaller organisms should be lower. However, there are unexplained differences that the equation cannot fully explain, like heterogeneity in resting metabolic rate of teleost fish (Clarke, 2006; Clarke \& Johnston, 1999). Others highlight that MTE makes valid assumptions and is more flexible than opponents argue, as it provides a foundational framework for general estimates of metabolic rate for larger, more generalizable taxonomic groups (Gillooly et al., 2006). Regardless of the debates about the details of MTE, at the very least, it provides a great foundation for thinking about temperature dependent processes.

MTE relies upon temperature impacting metabolic rate, as chemical reactions proceed faster at higher temperatures, which can have downstream effects on fitness characteristics. This is typically described by a characteristic unimodal, left skewed response that organisms typically demonstrate called a thermal performance curve (TPC) (Figure 1.2).
\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{figures/figure2}

@@

\caption[Thermal performance curve components]{Schematic of thermal performance curve and its important components for understanding thermal performance. CTmin denotes Critical Thermal Minimum, CTmax denotes Critical Thermal Maximum, and Topt denotes Thermal Optimum. Adapted from (Krenek et al 2012, Tuff et al 2016).}\label{fig:unnamed-chunk-4}
\end{figure}
The critical points that have been identified in thermal performance curves include: Thermal Optimum (Topt), Critical Thermal Minimum (CTmin), Critical Thermal Maximum (CTmax), Thermal tolerance breadth, and the Tolerance range. CTmin and CTmax describe the absolute limits of an organism's tolerance range, such that beyond these two thresholds, an organism cannot survive (Brown, Gillooly, Allen, Savage, \& West, 2004; Dowd, King, \& Denny, 2015). Topt is the temperature at which performance is highest, and is importantly situated just before the sharp decline in performance that converges on CTmax.

TPCs non-linear shape represents the relationship between temperature and performance, such that performance is an amalgamation of many life history characteristics and biological processes (e.g.~metabolism) (Amarasekare \& Savage, 2012). If only temperature's effect on metabolism were considered, the shape of the relationship may be an exponential one up to a certain point, after which it would no longer increase and the organism would fail to survive. However, there are several other constraining factors (e.g.~energy acquisition, nutrient availability, etc.) that drive the characteristically nonlinear pattern of TPCs. TPCs represent the complexity of many interacting processes an organism experiences in the environment with respect to thermal dependence, cementing their place as an important metric to conceptualize and understand temperature dependence.

TPCs have been crucial in predicting and understanding how organisms respond to their environment (Bernhardt, Sunday, Thompson, \& O'Connor, 2018). However, TPCs usually consist of measurements taken across an array of constant temperatures (e.g.~15-35°C in 5°C increments, with organisms reared at those temperatures constantly for the duration of an experiment to estimate thermal performance). These performance measurements are useful for a baseline understanding of thermal performance, however, they are not reflective of the variable conditions organisms experience in their natural environments.

Given the characteristically nonlinear shape of TPCs, Jensen's inequality is important in understanding the limits of thermal performance estimated under constant conditions versus variable conditions. Jensen's inequality describes the phenomenon that environmental variation can predictably and significantly impact nonlinear biological processes such that the consequences cannot be aptly inferred from average environmental conditions (Ruel \& Ayres, 1999). For instance, Ruel and Ayres provide the development rate of ectotherms as an example to demonstrate Jensen's inequality (1999). The development rate of ectotherms has a characteristically nonlinear shape, such that responses increase at lower temperatures, become linear at intermediate temperatures, and decrease at higher temperatures (Ruel \& Ayres, 1999). With Jensen's inequality, we can predict that the effects of temperature variation will be positive at lower temperatures, neutral at intermediate temperatures, and negative at higher temperatures (Ruel \& Ayres, 1999). Given this, it is important to recognize that the constant temperatures used to generate TPCs are simply averages of environmental conditions during a moment in time. It is therefore reasonable to extrapolate that thermal performance under truly variable conditions may drastically differ and has been demonstrated by subsequent studies (Bernhardt et al., 2018; Khelifa et al., 2019).

Beyond accurately predicting how variable environmental conditions affect thermal performance, the bounds and critical points of TPCs are important for understanding how changes in environmental conditions may differentially impact different groups of organisms. For instance, Amarasekare et al (2012) demonstrated the importance of where Topt is located relative to CTmax, such that ectotherms with a closer range between those two critical points at different geographic ranges may be negatively impacted by shifts in temperature and variability patterns (Amarasekare \& Savage, 2012; Paaijmans et al., 2013). Ectotherms in warm regions with an already minimized distance between Topt and CTmax may risk rapid extinction because there is less of a buffer to withstand increases in environmental temperature (Amarasekare \& Savage, 2012).

Outside of the debate about the equations for determining metabolic rate, many of these equations rely on constant temperature conditions to predict metabolic rate. Variation in temperature is predicted to affect performance more significantly than increasing temperatures (Vasseur et al., 2014); thus, if metabolic rates scale across species, communities, and ecosystems, how might non-constant temperature alter predictions regarding performance? Will increased variation have differing effects on different levels of biological organization? Khelifa et al (2019) demonstrated in five species from the same genus that when correcting for non-linearity, lab controlled experiments of diurnal variation in temperature and constant temperature did not differ significantly from one another in their thermal performance.

However, beyond diurnal or sinusoidal variation, there are other variation patterns at different timescales that all levels of biological organization experience (Figure 1.3). Khelifa et al (2019) demonstrated that when organisms are exposed to ambient temperatures in the environment, their thermal performance becomes less predictable. Beyond diurnal fluctuations in temperature, organisms can experience seasonal fluctuations as well as annual fluctuations in temperature. Vasseur and Yodzis outline how the frequency (1/period) of environmental fluctuations describes the ``color of environmental noise'' (Vasseur \& Yodzis, 2004). In colored noise, red noise describes lower frequencies and longer periods whereas white noise describes higher frequencies and shorter periods. Temperature is predicted to become more reddened or autocorrelated over time, meaning that temperatures will become more similar and predictable (Tabari, Hosseinzadeh Talaee, Ezani, \& Shifteh Some'e, 2012; Wigley, 1998). Understanding how organisms at all scales respond to these different patterns of variation in temperature is key for a more informed understanding of persistence.
\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{figures/figure4}

@@

\caption[Patterns of variability and spectral density plots]{Time series of daily changes in autocorrelated (A), stochastic (C), and diurnal (E) temperature variation over one year, accompanied by the respective spectral density plots for each pattern of variation (B,D,F). Adapted from and inspired by Massey et al 2020 and Kroeker et al 2020.}\label{fig:unnamed-chunk-5}
\end{figure}
\clearpage

Thermal fluctuations are expected to have an effect on performance across the individual, population, and community level:
\begin{itemize}
\item
  At the individual level, this usually comes in the form of reduced fitness, increased development rates, and/or decreased physiological responses from reduced maximum temperatures, from flattening the thermal performance curve (Bartheld, Artacho, \& Bacigalupe, 2017) to changes in TSD ratios (Bowden \& Paitz, 2018)
\item
  At the population level, this usually comes in the form of decreasing population densities (Anders \& Post, 2006)
\item
  At the community level, this usually comes in the form of promoting coexistence in the face of emigration (Descamps-Julien \& Gonzalez, 2005), or population success via inflationary effects (Holt et al., 2003)
\end{itemize}
It is, however, less understood how variation patterns influence these characteristic responses at each level of organization. Many individual level studies focus on responses to diurnal fluctuations in temperature, whereas studies at higher levels of biological organization have explored how colored noise elicits these characteristic responses (Petchey, 2000).

There has yet to be a quantitative review of the effect of environmental variability across ecological scales, even though many studies have looked at the significance of variability on ecosystem dynamics (Kroeker et al., 2020; Massey, Holt, Brooks, \& Rollinson, 2019). To understand how environmental variability affects organisms, populations, and communities, we conducted a meta-analysis on the effects of thermal variability on responses across different levels of biological organization to answer the following questions: 1) Does variability pattern drive differences in response magnitude? (i.e.~does autocorrelated variation have a larger impact on response than stochastic or diurnal variation?), 2) Does temperature variability drive differential response levels at different levels of biological organization?, 3) How can the magnitude of response be influenced by additional factors (e.g.~periodicity, age, size, thermal history, etc.)? We aim to answer these questions by collecting data from the literature at different levels of biological organization to systematically analyze responses across levels of organization.

We hypothesize that: 1) The larger the range of temperatures an organism will experience relative to their thermal performance curve will negatively impact the magnitude of response, 2) Additional covariates related to biological fitness and success, experimental age and organism size will be correlated with negative responses, such that older organisms will exhibit increased performance and smaller organisms will exhibit decreased performance with increased thermal variability, 3) Across organization levels, individual level studies will experience decreased performance compared to population level studies and higher levels of organization due to their ability to mitigate and rescue performance with population dynamics, 4) If organisms experience temperatures that exceed their Topt, their responses will be negatively impacted.

\hypertarget{methods}{%
\chapter{Methods}\label{methods}}
\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{figures/figure5}

@@

\caption[Meta-analysis paper selection process]{Diagram of the paper selection process from the original 176 papers returned from the Web of Science search to the 15 studies included in the analysis. Red indicates no usable data, magenta indicates some usable data but not for our study questions, and green indicates usable data flow for inclusion in analysis.}\label{fig:unnamed-chunk-7}
\end{figure}
We conducted a systematic literature search using the ISI Web of Science database to collect studies that investigated the effects of temperature variability (Figure 2.1). After trying several different combinations of search terms to capture research on temperature variability, our final search utilized these search terms in an advanced search: AK=((temperature OR thermal) NEAR (vari* OR fluc*)) AND SU=(Life Sciences \& Biomedicine). This search returned a total of 176 results. We downloaded the .bib file of this search and imported this file into the revtools R package (version 0.4.1). Using the revtools package we then screened studies by title, then by abstract, for inclusion in analysis (Figure 2.1). Once we screened studies for relevance to the field of biology and environmental variation, we began excluding studies using the following strict criteria:
\begin{itemize}
\tightlist
\item
  We excluded studies that did not feature both a constant and fluctuating treatment (e.g.~studies with several fluctuating treatments but no constant comparison or studies with a gradient of constant treatment temperatures but no periodic fluctuating treatment)
\item
  We excluded studies that were not completed in a laboratory (e.g.~modeling predictions or results from models, mesocosm experiments, \emph{in situ} experiments)
\item
  We excluded studies in which the responses were measured in different thermal conditions than the incubation or rearing thermal conditions (e.g.~organisms reared in fluctuating temperatures, but measured responses at different constant temperatures to yield a TPC)
\item
  We excluded any articles not written in English
\item
  We excluded studies that did not report measurements of error
\item
  We excluded studies in which the response variables were not comparable or extractable (e.g.~one study performed Principal Component Analysis for their responses)
\end{itemize}
We excluded studies that did not have the same mean for constant and fluctuating treatments (e.g.~treatments at the minimum and maximum with a fluctuation treatment across the range) or a mean within 1°C of the two treatments. We chose to exclude studies that did not match this criteria as there would not have been a baseline comparison to attribute changes in response magnitude most simply to variation, as changes in mean temperature significantly influence the magnitude of response (Brown et al., 2004). The final number of data points included in the analysis is broken down by study (Table 2.1).

\clearpage
\begin{table}[!h]

\caption[Studies included in analysis of thermal variability]{\label{tab:unnamed-chunk-8}Studies included in analysis of thermal variability (n=15)}
\centering
\begin{tabular}[t]{lll}
\toprule
\textbf{Study} & \textbf{Year} & \textbf{Extracted data points (n)}\\
\midrule
\cellcolor{gray!6}{Delava et al} & \cellcolor{gray!6}{2016} & \cellcolor{gray!6}{16}\\
Du et al & 2009 & 32\\
\cellcolor{gray!6}{Garcia-Ruiz et al} & \cellcolor{gray!6}{2011} & \cellcolor{gray!6}{16}\\
Glass et al & 2009 & 16\\
\cellcolor{gray!6}{Kern et al} & \cellcolor{gray!6}{2015} & \cellcolor{gray!6}{24}\\
\addlinespace
Kern et al & 2014 & 4\\
\cellcolor{gray!6}{Lowenborg et al} & \cellcolor{gray!6}{2012} & \cellcolor{gray!6}{10}\\
Maneti et al & 2014 & 24\\
\cellcolor{gray!6}{Pendlebury et al} & \cellcolor{gray!6}{2004} & \cellcolor{gray!6}{16}\\
Qu et al & 2014 & 40\\
\addlinespace
\cellcolor{gray!6}{Rolandi et al} & \cellcolor{gray!6}{2018} & \cellcolor{gray!6}{4}\\
Saxon et al & 2017 & 12\\
\cellcolor{gray!6}{Semenov et al} & \cellcolor{gray!6}{2007} & \cellcolor{gray!6}{32}\\
Simoncini et al & 2019 & 30\\
\cellcolor{gray!6}{Treidel et al} & \cellcolor{gray!6}{2015} & \cellcolor{gray!6}{4}\\
\bottomrule
\end{tabular}
\end{table}
\hypertarget{data-extraction-and-analysis}{%
\section{Data extraction and analysis}\label{data-extraction-and-analysis}}

From the studies that met our inclusion criteria, we extracted data from tables and figures. Figures were extracted via .png files imported into WebPlotDigitizer to obtain the raw data. The extracted data was then imported into a master spreadsheet of data from all studies for analysis. We extracted mean values for response variables, any measure of variance (Standard Deviation (SD) or Standard Error of the Mean (SEM)), and sample size. For studies that reported error as SEM, we converted SEM to SD by multiplying the SEM by the square root of the same SEM. We also collected information on mean temperature, fluctuation range, organization level, body size, life stage, response type, thermal stress, and aggregated these values (Table 2.2).

If studies reported findings using medians and the Interquartile Range (IQR) via boxplots, a normal distribution of the data was assumed and a subsequent calculation was performed, such that the median is the mean and the IQR is the SD with a conversion calculation (Higgins et al., 2011). Lastly, if extracted values were missing variance or sample size estimates, the points were automatically excluded via the data analysis software, \emph{metafor}.

\clearpage
\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-9}Covariates included in models}
\centering
\begin{tabular}[t]{l>{\raggedright\arraybackslash}p{8cm}}
\toprule
\textbf{Variable} & \textbf{Description}\\
\midrule
\cellcolor{gray!6}{mean\_temp\_constant} & \cellcolor{gray!6}{Reported mean temperature (°C) in study}\\
flux\_range & Reported range (difference between minimum 
                                     and maximum) of temperature fluctuation (°C) experienced\\
org\_level & Level of biological organization, categorized
                                     as either 0 for individual, 1 for population level
\cellcolor{gray!6}{                                     responses}\\
size & Size estimate of study organism, 0 for extra-small (less than 1mm),
                                     1 for small (less than 5 cm), 2 for medium (less than 0.3 m), and 3
                                     for large (greater than 0.3 m)\\
exp\_age & Life stage at which experiment was conducted, 
                                     categorized as 0 for larval, 1 for juvenile (not sexually mature),
\cellcolor{gray!6}{                                     and 2 for adult}\\
\addlinespace
resp\_type & Response variable units classified as rates or traits, 
                                     such that rates included any response variable units with a
                                     “per day” clarifier (e.g. growth rate), all else as traits (e.g. body
                                     width)\\
stressful & Whether or not the organisms experienced thermally stressful 
                                     temperatures throughout study duration, if temperature range 
                                     exceeded Topt, classified as thermally stressful, else not
                                     thermally stressful. See Appendix Table 1 for references in which
\cellcolor{gray!6}{                                     organisms’ thermal stress metrics were included}\\
\bottomrule
\end{tabular}
\end{table}
\clearpage

We used \emph{metafor} (version 2.4.0) in R to analyze the extracted, aggregated data. \emph{Metafor} builds on the functionality of older meta-analysis packages (meta, etc.) by allowing for greater flexibility in model types (mixed and random effects models) (Viechtbauer, 2010). In order to use \emph{metafor}, the package requires the data to be in a wider format, in which data are input into the following columns:
\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-10}Column names expanded to wider format for data analysis in metafor}
\centering
\begin{tabular}[t]{ll}
\toprule
\textbf{Column name} & \textbf{Purpose}\\
\midrule
\cellcolor{gray!6}{Mf} & \cellcolor{gray!6}{Mean of experimental group}\\
SDf & Standard deviation of the experimental group\\
\cellcolor{gray!6}{Mc} & \cellcolor{gray!6}{Mean of the control group}\\
SDc & Standard deviation of the control group\\
\cellcolor{gray!6}{Nf} & \cellcolor{gray!6}{Sample size (n) of the experimental group}\\
\addlinespace
Nc & Sample size (n) of the control group\\
\bottomrule
\end{tabular}
\end{table}
\linebreak

>>>>

% Syntax highlighting #22

@@

@@

@@

@@

Placeholder

\hypertarget{environmental-variability}{%
\section{Environmental variability}\label{environmental-variability}}

@@

@@

@@

@@

@@

@@

@@

\hypertarget{methods}{%
\chapter{Methods}\label{methods}}

Placeholder

\hypertarget{data-extraction-and-analysis}{%
\section{Data extraction and analysis}\label{data-extraction-and-analysis}}

\hypertarget{results}{%
\chapter{Results}\label{results}}
\begin{figure}

{\centering \includegraphics[width=0.95\linewidth]{thesis_files/figure-latex/unnamed-chunk-2-1}

@@

\caption[Forest plot of effect sizes]{Forest plot of effects sizes by study and experiment within a study. Data indicate that both within and across studies, there were a wide range of effect sizes, likely due to experimental design (including fluctuation range, mean temperature, etc.) Each point represents an average of SMD for multiple responses within each experiment in a single study. Accordingly, the error bars represent an average of variance across multiple responses within a single experiment.}\label{fig:unnamed-chunk-2}
\end{figure}
Overall, even our simple model demonstrated that the intercept was not significant but there was significant heterogeneity across studies (n=140, z= 0.79, p \textless{} 0.0001), as demonstrated by Figure 3.1. These results allowed us to reject the null hypothesis that there are no significant differences between SMD across the studies included in analysis (see Appendix Figure 1). This then allowed us to proceed in further detail how different covariates were influencing the effect sizes specifically.

{\centering \includegraphics[width=0.95\linewidth]{thesis_files/figure-latex/unnamed-chunk-3-1}

@@

\caption[Coefficient estimates for full model]{Coefficient estimates for each covariate in the full model, where * signify statistically significant model predictors.  Error bars represent 95 percent confidence intervals. Yellow indicates a positive coefficient estimate, purple indicates a negative coefficient estimate.}\label{fig:unnamed-chunk-3}
\end{figure}
\clearpage

\caption[Thermal stress model summary statistics]{\label{tab:unnamed-chunk-4}Model summary statistics for random effects model using thermal stress as sole modifier}
\centering
\begin{tabular}[t]{lrrrrrr}
\toprule
\textbf{Term} & \textbf{estimate} & \textbf{se} & \textbf{zval} & \textbf{pval} & \textbf{ci.lb} & \textbf{ci.ub}\\
\midrule
\cellcolor{gray!6}{intercept} & \cellcolor{gray!6}{0.2050} & \cellcolor{gray!6}{0.1846} & \cellcolor{gray!6}{1.1104} & \cellcolor{gray!6}{0.2668} & \cellcolor{gray!6}{-0.1568} & \cellcolor{gray!6}{0.5669}\\
stressful & -0.0489 & 0.0702 & -0.6975 & 0.4855 & -0.1864 & 0.0886\\
\bottomrule
\end{tabular}
\end{table}
\clearpage

{\centering \includegraphics[width=0.9\linewidth]{thesis_files/figure-latex/unnamed-chunk-5-1}

@@

\caption[Scatterplot of relationship between range, mean, and SMD]{Scatterplot describing the relationship between mean temperature, temperature fluctuation range, and SMD. Temperature fluctuation ranged from 0-20°C, mean temperature ranged from 7-33°C. SMD ranged from -4 to 6, restricted based on the distribution of SMD to minimize impact of outliers (see Appendix Figure 2).}\label{fig:unnamed-chunk-5}
\end{figure}
The trend between organization level and temperature fluctuation range, suggests that populations perform better at higher fluctuation ranges than organisms (Figure 3.4). Interestingly, there were not any significant differences between thermal stress (Figure 3.5), life history metric (Figure 3.6), body size (Figure 3.7), or age (Figure 3.8).

{\centering \includegraphics[width=0.85\linewidth]{thesis_files/figure-latex/unnamed-chunk-6-1}
